# 1st Tutorial  


::: {.exercise}
Describe briefly each of the following terms:
:::

_a. Primary data_
\
The data collected by the investigator from the original source.
\
\
*b. Secondary data*
\
If the required data had already been collected by some agencies or individuals and are now available in the published or unpublished records.
\
\
*c. Mail inquiry*
\
The investigator prepares a questionnaire and sends it by mail to the respondents.
\
\

*d. Questionnaire/schedule*
\
The questionnaire is a channel through which the needed information is elicited.
\
\
*e. Population*
\
The population or universe is an aggregate of elements, about which the inference is to be made.
\
\
*f. Census*
\
Collection of information on every unit in the population for the characteristics of interest. (Also known as enumeration).
\
\
*g. Element*
\
An element is a unit for which information is sought.
\
\
*h. Sample*
\
A subset of population selected from a frame to draw inferences about a population characteristic.
\
\
*i. Sampling unit*
\
Non-overlapping collections of elements of the population.
\
\
*j. Sampling frame*
\
A list of all the units in the population to be sampled is termed


::: {.exercise}
Describe briefly the difference between with and without replacement sampling.
:::
\
the units that selected one by one from the population, in WR the unit selected at any particular draw is replaced back to the population before selecting a unit at the next draw, but in the WOR is not replaced.

::: {.exercise}
Consider a population consisting of 6 villages, the areas (in hectares) of which are given below
:::
```{r , echo=FALSE}
frame = data.frame(Village=c("A","B","C","D","E","F"),Area = c(760,343,657,550,480,935))
frame=t(frame)
knitr::kable(frame, raw.names = c("X","Y"), align = "lccrr")
```

a. Enumerate all possible WR samples of size 3. Also, write the values of the study variable for the sampled units.
\
\

$$N=6 ~~~~~~~~~~~~~~~~~ n=3$$
Thus the possible samples $= N^n = 6^3 = 216$
\

```{r , echo=FALSE}
frame = data.frame(Village=c("AAA","AAB","ABA","BAA","AAC","....."),Area = c("760 760 760","760 760 343","760 343 760","343 760 760","760 760 657","    "))

knitr::kable(frame, col.names = c("Villages in the sample","Area of the villages"), align = "lccrr")
```



b. List all the WOR samples of size 4 along with their area values.
\
The number of possible samples $= C_n^N= C_4^6=15$
\
All possible WR samples along with area values are listed below:

```{r, echo=FALSE}
DF = read.csv("DATA.csv", header=TRUE)
knitr::kable(DF, col.names = c("Sample" ,"Villages in the sample","Area of the villages"), align = "lccrr")

```


\
\ 


